Kernel Square-Loss Examplar Machines

### 2016-11-07 esvm_compversion.tex
first commit

\usepackage{times}
\usepackage{bbm}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{caption}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{coro}{Corollary}
\newtheorem{prop}{Property}
\newtheorem{definition}{Definition}

@@

\section{Square-loss examplar machines\label{sec:sqesvm}}
Given some loss function $L:[-1,+1]\times \RR\rightarrow \RR^+$, some
positive training example $x_0$ in $\RR^p$ and $n$ negative training
examples $x_i$ in $\RR^p$ ($i=1,\ldots,n$), training an exemplar
linear classifier consists in minimizing the error
\begin{equation}
E(w,b)=\frac{1}{n}\sum_{i=1}^n L(-1,w\cdot x_i+b)+ L(1,w\cdot
x_0+b)+\frac{\lambda}{2} ||w||^2
\label{eq:general}
\end{equation}
with respect to $w$ in $\RR^p$ and $b$ in $\RR$.

The traditional examplar SVM uses the hinge loss for $L$. Let us use
instead the square loss. Equation~(\ref{eq:general}) becomes
\begin{equation}
E(w,b)=\frac{1}{2n}\sum_{i=1}^n (w\cdot x_i+b+1)^2+ \frac{1}{2}(w\cdot
x_0+b-1)^2+\frac{\lambda}{2} ||w||^2.
\label{eq:square}
\end{equation}

The interest of the square loss is that the minimization can be done
in closed form. Indeed, writing that $\partial E/\partial b=0$ at a
minimum yields
\begin{equation}
b=-\frac{1}{2}(x_0+\mu)\cdot w,
\end{equation}
where $\mu=\frac{1}{n}\sum_{i=1}^n x_i$ is the center of mass of the
negative samples $x_i$.

Substituting in Eq.~(\ref{eq:square}) now reduces our optimization problem
to the minimization with respect to $w$ of
\begin{equation}
\begin{array}{lcl}\hat{E}(w)&=&
\displaystyle\frac{1}{2n}\sum_{i=1}^n [w\cdot (x_i-\frac{1}{2}(x_0+\mu))+1]^2
+\frac{1}{2}[w\cdot (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||w||^2\\
&=&
\displaystyle\min_{w\in\RR^p}
\frac{1}{2n}\sum_{i=1}^n [w\cdot (\sigma_i-\frac{1}{2}\delta) +1]^2
+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)^2+\frac{\lambda}{2}||w||^2,
\end{array}
\label{eq:weq}
\end{equation}
where $\delta=x_0-\mu$ and $\sigma_i=x_i-\mu$.

Finally, writing $\partial\hat{E}/\partial w=0$ at a minimum yields
\begin{equation}
\frac{1}{n}\sum_{i=1}^n [w\cdot(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)\delta +\lambda w=0,
\end{equation}
or
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] w
+\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T w -\frac{1}{2}\delta +\lambda w
=0.
\end{equation}

Using the fact that the variables $\sigma_i$ have zero
mean, we obtain
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n\sigma_i\sigma_i^T +\frac{1}{2}\delta\delta^T +\lambda\text{Id}]
w=\delta,
\end{equation}
or, putting everything together and reverting to the original notation,
\begin{equation}
\left\{\begin{array}{l}
\displaystyle w=A^{-1} (x_0-\mu),\quad\text{where}\quad
A=\Sigma
+\frac{1}{2}(x_0-\mu)(x_0-\mu)^T +\lambda\text{Id},\\
\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot w,
\end{array}\right.
\label{eq:final}
\end{equation}
where
\begin{equation}
\Sigma=\frac{1}{n}\sum_{i=1}^n(x_i-\mu)(x_i-\mu)^T
=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}XX^T-\mu\mu^T,
\label{eq:Sigmadef}
\end{equation}
is the covariance matrix of the negative samples $x_i$, and
$X$ denotes the $p\times n$ matrix with columns $x_i^T$ ($i=1,\ldots,n$).

The vector $w$ (or the vector $[w;b]$) of this {\em square-loss
examplar machine} can potentially be used as a new representation for
the corresponding positive feature $x_0$. Note that this gives
asymmetric roles to positive and negative examples.

@@

\section{Kernelized version}
Let us recall a few basic facts avout kernel methods for supervised
classification. We consider a reprodcing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel. We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n L[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
\label{eq:kernel}
\end{equation}
where the pairs $(x_i,y_i)$ ($i=1,\ldots,n$) in $X\times Y$ are
training samples, $Y=\RR$ (regression) or $Y=\{-1,1\}$
(classification), and $L: Y\times \RR\rightarrow\RR$ is some arbitrary
loss function. By definition of a reproducing kernel,
Eq.~(\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n L[y_i,( \varphi(x_i)|h)+b] +
\frac{\lambda}{2}||h||^2,
\label{eq:kernelaff}
\end{equation}
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form). We dub
problems with the general form of (\ref{eq:kernelaff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $b$, $(h|h')+b$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

@@

The primal problem is
\begin{equation}
\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n L[y_i,k_i^T\alpha+b]
+\frac{\lambda}{2}\alpha^TK\alpha,
\label{eq:primal1}
\end{equation}
and any solution $(\alpha^\star,b^\star)$ to (\ref{eq:primal1})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.

Let $r$ denote the rank of the kernel matrix $K$, a
{\em reduced} version of the primal problem is
\begin{equation}
\min_{\beta\in\RR^r,b\in\RR} \sum_{i=1}^n L[y_i,b_i^T\beta+b]
+\frac{\lambda}{2}||\beta||^2,
\label{eq:primal2}
\end{equation}
where $B$ is the $n\times r$ square root of $K$ and $b_i^T$ denotes
its $i$th row,
and any solution $(\beta^\star,b^\star)$ to (\ref{eq:primal2})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$, and
$\alpha^\star=P\beta^\star$, where $P$ is the pseudo-inverse of $B^T$.

The dual problem is
\begin{equation}
\min_{\alpha\in\RR^n}
\sum_{i=1}^n L^*(y_i,-\lambda\alpha_i)
+\frac{\lambda}{2}\alpha^TK\alpha
\quad\text{subject to}\quad \sum_{i=1}^n \alpha_i=0,
\label{eq:finaldual}
\end{equation}
and the optimal value $b^*$ of $b$ given any solution
$\alpha^\star$ of (\ref{eq:finaldual}) can be found by solving
\begin{equation}
\min_{b\in \RR} \sum_{i=1}^n L(y_i,k_i^T\alpha^{\star}+b).
\label{eq:minb}
\end{equation}
The pair $(\alpha^\star,b^\star)$ then provides a solution
$(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with $h^\star=
\sum_{i=1}^n\alpha_i^\star\varphi(x_i)$.
\end{theorem}

@@

Theorem\ref{th:affinelearn} follows fromt the representer theorem
~\cite{SHS01,Wahba90} and elementary results from analysis and
optimization theory.

@@

The next section shows how to train the kernel square-loss examplar
machine efficiently.

\section{Efficient implementation}
Let us consider the $n\times n$ kernel matrix $\hat{K}$ associated with
the $n$ negative examples, and a low-rank factorization $\hat{K}=\hat{B}\hat{B}^T$
of
this matrix, where $\hat{B}$ is an
$n\times r$ square root of $\hat{K}$. This factorization can be computed in $O(nr^2)$ time
and $O(nr)$ storage~\cite{BaJo05,FiSc01}. The singular value
decomposition $UWV^T$ of $\hat{B}$ and its $r\times n$ pseudoinverse
$\hat{B}^\dagger=(\hat{B}^T\hat{B})^{-1}\hat{B}^T=VW^{-1}U^T$ such that $\hat{B}^\dagger
\hat{B}=\text{Id}$ can also be computed with the same time and storage
costs.

Let us consider the augmented
kernel matrix $K$ obtained by adding one more (positive) sample. This matrix
can be written as
\begin{equation}
K=\qmatrix{k_{00} & k_0^T\\k_0& \hat{K}},
\end{equation}
where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
The following
lemma show that a $(n+1)\times (r+1)$ factorization of $K$ can
be computed efficiently as well.

\begin{lemma}
The augmented kernel matrix $\hat{K}$ can be factorized as
\begin{equation}
\hat{K}=BB^T,\quad\text{where}\quad
B=\qmatrix{u & v^T\\0 & \hat{B}}\quad\text{and}\quad
v=\hat{B}^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}
\end{equation}
in time $O(nr)$ and storage $O(n)$.
\end{lemma}
\begin{proof}
Indeed,
\begin{equation}
BB^T=\qmatrix{u^2+||v||^2 & v^T\hat{B}^T\\ \hat{B}v & \hat{B}\hat{B}^T}.
\end{equation}
Since $K$ is positive semidefinite $k_0$ must lie in the column space
$\cal \hat{B}$ of $\hat{B}$,\footnote{Let
us assume that $k_0$ does not belong to $\cal \hat{B}$, then it can be
decomposed uniquely as $k_0=l_0+m_0$, where $l_0$ belongs to $\cal \hat{B}$,
and $m_0\neq 0$ belong to the orthogonal complement $\cal \hat{B}^\perp$ of
$\cal \hat{B}$. Obviously $(1,-\nu m_0^T)K(1;-\nu m_0)=k_{00}-2\nu||m_0||^2$, which
is negative for large enough values of $\nu>0$. In turn, this contradicts the
fact that a kernel matrix is positive semidefinite.}
and $b=\hat{B}^\dagger k_0$ is thus an exact solution of
$\hat{B}b=k_0$.
The fact that $k_{00}-||b||^2$ is nonnegative follows immediately from
the fact that the Schur complement $\hat{K}-||k_0||^2/k_{00}$ of $k_{00}$ in
the positive semidefinite matrix $K$ is itself positive semidefinite.\footnote{Indeed,
since the matrix $k_{00}K-k_0k_0^T=\hat{B}(k_{00}\text{Id}-bb^T)\hat{B}^T$ is
positive semidefinite, and $\hat{B}$ has rank $r$, the $r\times r$ matrix
$k_{00}\text{Id}-bb^T$
is also positive definite, thus
$b^T(k_{00}\text{Id}-bb^T)b=||b||^2(k_{00}-||b||^2)\ge 0$.} The
cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
of $b$
\end{proof}

Training the kernel square-loss examplar machine given $\hat{B}$ amounts to
computing
$\Sigma$ in (\ref{eq:Sigmadef}) and solving the $(r+1)\times (r+1)$
linear system in (\ref{eq:final}). Note that, because of the
particular structure of $B$, only one term in the sum in (\ref{eq:Sigmadef})
must be recomputed for each new positive example, with $O(r^2)$ cost.
The linear system itself can be solved in $O(r^3)$ cost.

Once the optimal parameters $\beta^\star$ haven been found, they can
be used directly as some representation of the positive example. The
corresponding value $\alpha^\star$ (or more correctly, {\em a}
corresponding value) can be computed as $\alpha^\star =P\beta^\star$,
where $P=B(B^TB)^{-1}$ is the pseudoinverse of $B^T$ with
$B^TP=\text{Id}$. The corresponding element of $\cal H$ is
$h^\star=\sum_{i=0}^n \alpha_i^\star\varphi(x_i)$.

Now, given two positive images $x_0$ and $x_0'$, they can be compared
by computing the normalized inner product
\begin{equation}
s(x_0,x_0')=\frac{(h|h')}{\sqrt{(h|h)}\sqrt{(h'|h')}}
\end{equation}
of the corresponding elements $h$ and $h'$ of $H$ as a similarity measure
(we drop the ``$\star$'' exponents from now on to avoid clutter in the
noration).

Concretely, we have
\begin{equation}
\begin{array}{lcl}
(h|h')&=&\displaystyle[\alpha_0\varphi(x_0)+\sum_{i=1}^n\alpha_i\varphi(x_i)
|\alpha_0'\varphi(x_0')+\sum_{i=1}^n\alpha_i'\varphi(x_i)]\\
&=&\displaystyle
\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0\sum_{i=1}^n\alpha_i' k(x_0,x_i)
+\alpha_0'\sum_{i=1}^n
\alpha_ik(x_0',x_i)+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
&=&\displaystyle\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0k_0^T\hat{\alpha}'
+\alpha_0'k_0'^T\hat{\alpha}
+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
\end{array}
\label{eq:lcl}
\end{equation}
where $\hat{\alpha}$ (resp. $\hat{\alpha}'$) denotes the vector of
$\RR^n$ formed by excluding the first (or rather zeroth) element
of $\alpha$ (resp. $\alpha'$).

If we write $\beta = [\beta_0; \hat{\beta}]$ similarly to
$\alpha=[\alpha_0; \hat{\alpha}]$ and keeping in mind that
$P=[\frac{1}{u}\ 0; -\frac{1}{u}\hat{P}v \ \hat{P}]$, where $\hat{P}$
is the pseudoinverse of
$\hat{B}^T$, we have that
\begin{equation}
\alpha_0 = \dfrac{\beta_0}{u}, \,\, \hat{\alpha}=-\dfrac{\beta_0}{u}Pv+P\hat{\beta}.
\end{equation}
Remembering that $\hat{K}=\hat{B}\hat{B}^T$ and and $k_0=\hat{B}v$
allows us to rewrite (\ref{eq:lcl}) as
\begin{align}
(h|h')& = \dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
+\dfrac{\beta_0}{u}v^T\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)
+\dfrac{\beta_0'}{u'}v'^T\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\\ &+\left(\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\right)^T
\left(\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)\right)\\ &=
\dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
+\dfrac{\beta_0}{u}v^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)
+\dfrac{\beta_0'}{u'}v'^T\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)
+
\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)\\ &=
\hat{\beta}^T\hat{\beta'}
+\dfrac{\beta_0\beta_0'}{uu'}\left(k(x_0,x_0')-v^Tv'\right).
\end{align}
We also note that $(h|h)=||\beta||^2$ and $(h'|h')=||\beta'||^2$.

>>>>

\usepackage{times}
\usepackage{bbm}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{caption}
\usepackage{subcaption}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{coro}{Corollary}
\newtheorem{prop}{Property}
\newtheorem{definition}{Definition}

@@

\section{Square-loss exemplar machines\label{sec:sqesvm}}
Given some loss function $l:[-1,+1]\times \RR\rightarrow \RR^+$, some
positive training example $x_0$ in $\RR^p$ and $n$ negative training
examples $x_i$ in $\RR^p$ ($i=1,\ldots,n$), training an exemplar
linear classifier consists in minimizing the error
\begin{equation} 
E(\omega,b)=\frac{1}{n}\sum_{i=1}^n l(-1,\omega^T x_i+b)+ \theta \ l(1,\omega^T x_0+b)+\frac{\lambda}{2} ||\omega||^2
\label{eq:general}
\end{equation} 
with respect to $\omega$ in $\RR^p$ and $b$ in $\RR$.

The traditional exemplar SVM uses the hinge loss for $l$. Let us use
instead the square loss. Equation~(\ref{eq:general}) becomes
\begin{equation}
E(\omega,b)=\frac{1}{2n}\sum_{i=1}^n (\omega^T x_i+b+1)^2+ \frac{\theta}{2}(\omega^T
x_0+b-1)^2+\frac{\lambda}{2} ||\omega||^2.
\label{eq:square}
\end{equation}

The interest of the square loss is that the minimization can be done
in closed form. Indeed, writing that $\partial E/\partial b=0$ at a
minimum yields
\begin{equation}
b=\dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta +1}(\theta x_0+\mu)^T \omega,
\end{equation}
where $\mu=\frac{1}{n}\sum_{i=1}^n x_i$ is the center of mass of the
negative samples $x_i$.

Substituting in Eq.~(\ref{eq:square}) now reduces our optimization problem
to the minimization with respect to $\omega$ of
\begin{equation}
\begin{array}{lcl}\hat{E}(\omega)&=&
\displaystyle\frac{1}{2n}\sum_{i=1}^n [\omega^T (x_i-\frac{1}{2}(x_0+\mu))+1]^2
+\frac{\theta}{2}[\omega^T (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||\omega||^2\\
&=&
\displaystyle\min_{\omega\in\RR^p}
\frac{1}{2n}\sum_{i=1}^n [\omega^T (\sigma_i-\frac{1}{2}\delta) +1]^2
+\frac{\theta}{2}(\frac{1}{2}\omega^T\delta-1)^2+\frac{\lambda}{2}||\omega||^2,
\end{array}
\label{eq:weq}
\end{equation}
where $\delta=x_0-\mu$ and $\sigma_i=x_i-\mu$.

Finally, writing $\partial\hat{E}/\partial \omega=0$ at a minimum yields
\begin{equation}
\frac{1}{n}\sum_{i=1}^n [\omega^T(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
+\frac{1}{2}(\frac{1}{2}\omega\cdot\delta-1)\delta +\lambda \omega=0,
\end{equation}
or
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] \omega
+\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T \omega -\frac{1}{2}\delta +\lambda \omega
=0.
\end{equation}

Using the fact that the variables $\sigma_i$ have zero
mean, we obtain
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n\sigma_i\sigma_i^T +\frac{1}{2}\delta\delta^T +\lambda\text{Id}]
\omega=\delta,
\end{equation}
or, putting everything together and reverting to the original notation,
\begin{equation}
\left\{\begin{array}{l}
\displaystyle \omega=U^{-1} (x_0-\mu),\quad\text{where}\quad
U=\Sigma
+\frac{1}{2}(x_0-\mu)(x_0-\mu)^T +\lambda\text{Id},\\
\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot \omega,
\end{array}\right.
\label{eq:final}
\end{equation} 
where 
\begin{equation}
\Sigma=\frac{1}{n}\sum_{i=1}^n(x_i-\mu)(x_i-\mu)^T
=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}X^TX-\mu\mu^T,
\label{eq:Sigmadef}
\end{equation}
is the covariance matrix of the negative samples $x_i$, and
$X$ denotes the $n\times p$ matrix with rows $x_i^T$ ($i=1,\ldots,n$).

The vector $\omega$ (or the vector $[\omega;b]$) of this {\em square-loss
examplar machine} can potentially be used as a new representation for
the corresponding positive feature $x_0$. Note that this gives
asymmetric roles to positive and negative examples.

@@

\section{Kernelized version}
Let us recall a few basic facts about kernel methods for supervised
classification. We consider a reproducing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel.  We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n l[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
\label{eq:kernel}
\end{equation}  
where the pairs $(x_i,y_i)$ ($i=1,\ldots,n$) in $X\times Y$ are
training samples, $Y=\RR$ (regression) or $Y=\{-1,1\}$
(classification), and $L: Y\times \RR\rightarrow\RR$ is some arbitrary
loss function. By definition of a reproducing kernel,
Eq.~(\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n l[y_i,( \varphi(x_i)|h)+b] +
\frac{\lambda}{2}||h||^2,
\label{eq:kernelaff}
\end{equation} 
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form). We dub
problems with the general form of (\ref{eq:kernelaff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $b$, $(h|h')+b$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

@@

The primal problem is
\begin{equation}
\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n l[y_i,k_i^T\alpha+b]
+\frac{\lambda}{2}\alpha^TK\alpha,
\label{eq:primal1}
\end{equation}
and any solution $(\alpha^\star,b^\star)$ to (\ref{eq:primal1})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.

Let $r$ denote the rank of the kernel matrix $K$, a 
{\em reduced} version of the primal problem is 
\begin{equation}
\min_{\beta\in\RR^r,b\in\RR}  \sum_{i=1}^n l[y_i,b_i^T\beta+b]
+\frac{\lambda}{2}||\beta||^2,
\label{eq:primal2}
\end{equation}
where $B$ is the $n\times r$ incomplete Cholesky decomposition of $K$ and $b_i^T$ denotes
its $i$th row,
and any solution $(\beta^\star,b^\star)$ to (\ref{eq:primal2})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$, and
$\alpha^\star=P\beta^\star$, where $P$ is the pseudo-inverse of $B^T$.

The dual problem is
\begin{equation}
\min_{\alpha\in\RR^n}
\sum_{i=1}^n l^*(y_i,-\lambda\alpha_i) 
+\frac{\lambda}{2}\alpha^TK\alpha
\quad\text{subject to}\quad \sum_{i=1}^n \alpha_i=0,
\label{eq:finaldual}
\end{equation}
and the optimal value $b^*$ of $b$ given any solution
$\alpha^\star$ of (\ref{eq:finaldual}) can be found by solving
\begin{equation}
\min_{b\in \RR} \sum_{i=1}^n l(y_i,k_i^T\alpha^{\star}+b).
\label{eq:minb}
\end{equation}
The pair $(\alpha^\star,b^\star)$ then provides a solution
$(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with $h^\star=
\sum_{i=1}^n\alpha_i^\star\varphi(x_i)$.
\end{theorem}

@@

Theorem\ref{th:affinelearn} follows from the representer theorem
~\cite{SHS01,Wahba90} and elementary results from analysis and
optimization theory.

@@

Once the optimal parameters $\beta^\star$ haven been found, they can
be used directly as some representation of the positive example.  The
corresponding value $\alpha^\star$ (or more correctly, {\em a}
corresponding value) can be computed as $\alpha^\star =P\beta^\star$,
where $P=B(B^TB)^{-1}$ is the pseudoinverse of $B^T$ with
$B^TP=\text{Id}$. The corresponding element of $\cal H$ is
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.  It can be compared
to any other element $h'=\sum_i^n\alpha_i'\varphi(x_i')$ of $\cal H$ (say, the representation of
another image) by computing $(h^\star|h')=\alpha^{\star T}K \alpha' = \beta^{\star T}\beta'$.

\section{Incomplete Cholesky decomposition}

In this section we revisit how to obtain a low-rank decomposition $B$ that better approximates $K$, as seen in ~\cite{BaJo05,FiSc01}.
Let us assume a sequence of indexes $I=\{i_1,i_2,\dots ,i_r\}$ from $\{1,2,\dots, n\}$ is given, where $r$ is the number of columns of $B$. 
$I$ is the set of pivots, chosen such that to minimize the error approximation between $BB^T$ and $K$. We also note, for $t$ in $\{1,2,...,r\}$, $I_t\{i_1,...,i_t\}$, $J_t=\{1,2,...,n\}/ I_t$ and $J=J_r$.

Given $I$ and $J$, we can calculate the $t$-th column, $t=1,2,...,r$, of $B$ interactively:
\begin{align}
\begin{cases}
\vspace{3 mm}
B(i_t, t) &= \left(K(i_t,i_t)-\displaystyle{\sum_{m=1}^{t-1}} B(i_{m},m)  \right)^{\frac{1}{2}},\\
\vspace{3 mm}
B(I_{t-1},t) &= 0,\\
B(J_t, t) &= \dfrac{1}{B(i_t,t)}\left(K(J_t,i_t)-\displaystyle{\sum_{m=1}^{t-1}B(J_t,m)B(i_t,m)}\right).
\end{cases}\label{icd:algo}
\end{align}
The time complexity of the $t$-th interaction is $O(tn)$, thus making the full algorithm $O(nr^2)$ in time. Since $B$ is $n\times r$, storage complexity is $O(nr)$.
\begin{proposition}
There is an unique $n\times n$ matrix $L$ such that
\begin{itemize}
\item L is symmetric and positive semidefinite,
\item L(:,I) = K(:,I),
\item the column space of $L$ is equal to the column space of $L(:,I)$.
\end{itemize}
L is such that 
\begin{equation}
L([I \ J],[I\ J]) = \qmatrix{K(I,I) & K(J,I)^T\\ K(J,I) & K(J,I)K(I,I)^{-1}K(J,I)^T}.
\end{equation}
\end{proposition}
From \textbf{Proposition 1}, $L$ is an unique approximation of $K$ for a given pair $I$ and $J$. The algorithm from (\ref{icd:algo}) construct $B$ such that $L=BB^T$.

The next section shows how to train the kernel square-loss exemplar machine efficiently.

\section{Efficient implementation of kernel decomposition}

Let us consider a low-rank factorization $K=BB^T$, where $B$ as given in the previous section. 
%The singular value decomposition $UWV^T$ of $B$ and its $r\times n$ pseudoinverse
%$B^\dagger=(B^TB)^{-1}B^T=VW^{-1}U^T$ such that $B^\dagger B=\text{Id}$ can also be computed with the same time and storage costs.

Let us consider the augmented
kernel matrix  $K'$ obtained by adding one more sample. This matrix
can be written as
\begin{equation}
K'=\qmatrix{k_{00} & k_0^T\\k_0& K},
\end{equation}
where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
The following
lemma shows that a $(n+1)\times (r+1)$ factorization of $K'$ can
be computed efficiently as well. In order to ease notations, we assume $[k_{00}; k_0]$ and $[k_{00}; k_0]^T$ are respectively the $0$-th column and row of $K'$. Hence, the $i$-th row of $K$ is part of the $i$-th row of $K'$, $i=\{1,2,...,n\}$. Also, let $I'=\{0\}\cup I$ be the set of pivots for the decomposition of $K'$.

\begin{lemma}
The augmented kernel matrix $K'$ can be factorized as
\begin{align}
K'&\approx B'B'^T,\quad\text{where}\quad
B'=\qmatrix{u & v^T\\w & B}\quad\text{and}\\
v=B_I^{-1} k_0(I),\,\, u&=\sqrt{k_{00}-||v||^2},\,\, w(I) = 0,\,\, w(J)=\dfrac{1}{u}(k_0(J)-B_Jv)
\end{align}
in time $O(nr)$ and storage $O(n)$. Here, $B_I$ and $B_J$ denote $B(I,:)$ and $B(J,:)$, respectively.
\end{lemma}
\begin{proof}
Let $L' = B'B'^T$. Indeed, 
\begin{equation}
L'=\qmatrix{u^2+||v||^2 & v^TB^T+uw^T\\ Bv+uw & BB^T+ww^T}. 
\end{equation}
Following \textbf{Proposition 1}, if $L'(I,:)=K'(I,:)$ and $L'(:,I)=K'(:,I)$, $w(I)=0$.
From the fact that $B$ has full rank, $B_I$ is a $r\times r$ matrix of rank $r$.
$B_I$ is therefore invertible, which guarantees the system $B_Iv=k_0(I)$ has a solution given by $v=B_I^{-1} k_0(I)$.
To show that $k_{00}\ge ||v||^2$, we first note the 
\begin{equation}
K'(I',I') = \qmatrix{k_{00} & k_0(I)^T\\ k_0(I)& K(I,I)}
\end{equation}
is positive semidefinite. Thus the Schur complement of $k_{00}$ is also non-negative:
\begin{equation}
k_{00}-k_0(I)^TK(I,I)^{-1}k_0(I)\ge 0,
\end{equation}
from which it follows that $k_{00}-||v||^2\ge 0$. \footnote{$k_0(I)^TK(I,I)^{-1}k_0(I)=k_0(I)^T(B_IB_I^T)^{-1}k_0(I)=v^Tv$}
$w(J)$ follows from $B_Jv+uw(J)=k_0(J).$

The cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
of $w$.
\end{proof}

\subsection{Why $w\ne 0$?}
In a previous version of this presentation, it argued for a decomposition of $K'$ with $w=0$.
One of the keys justifications of this assumption is that $B^\dagger v=k_0$ has a solution because $k_0$ lie in the column space $\cal B$ of $B$. 
This is not true however for a low-rank decomposition $B$, i.e. $r$ is smaller the rank of $K$. 
Indeed, $K'$ is a positive semidefinite matrix, but $[k_{00}\ k_0^T; k_0\ BB^T]$ is not. 
If $r$ is the rank of $K$, then we do find $w=0$.
But in practice, for $k_0$ does not belong in $\cal B$ and $k_{00}<||v||^2$ for $r$ not big enough.

\section{Efficient implementation of ridge regression}
Let us rewrite the kernelized problem after adding the positive exemplar to Equation (\ref{eq:primal2}):
\begin{equation}
\min_{\beta\in\RR^{r+1}, \ b\in\RR}\left( \theta (1-\beta^Tb_0'-b)^2+\dfrac{1}{n}\sum_{i=1}^n (1+\beta^Tb_i'+b)^2 +||\beta||^2\right).
\end{equation}
We recall the solution (\ref{eq:final}):
\begin{align}
\left\{\begin{array}{l}
\vspace{5 mm}
\displaystyle \beta^\star=\dfrac{2\theta}{\theta+1}U^{-1} \delta,\quad\text{where}\quad
U=\dfrac{1}{n}\hat{B}^T\hat{B}-\mu'\mu'^T
+\frac{\theta}{\theta+1}\delta\delta^T +\lambda\text{Id},\\
\displaystyle b^\star=\dfrac{\theta-1}{\theta+1}-\frac{1}{\theta+1}(b'_0+\mu')^T \omega,
\end{array}\right.
\label{kernel:final}
\end{align} 
Here, $b_i'^T$ is the $i$-th row of $B'$
\footnote{As for $K'$, we start counting at $0$ for $B'$ indexes.}, $\hat{B}$ is the matrix of rows $b_i'^T$, $i$ in $\{1,2,...,n\}$,
$\mu' = \frac{1}{n}\sum_{i=1}^nb'_i$ and $\delta = b'_0-\mu'$.
In parallel with the notations for $B$, we note as well $b_i^T$ is the $i$-th row of $B$, $\mu = \frac{1}{n}\sum_{i=1}^nb_i$.

We note $A$ the matrix $(n+1)\times (n+1)$ such that
\begin{equation}
A = \dfrac{1}{n}B'^TB'-\mu'\mu'^T+\lambda\text{Id}.\label{eq:A}
\end{equation}
